\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-06-21 paper.tex
save

Following \citep{2012arXiv1212.4507S}, variational optimization (VO) (also known as the search gradient algorithm~\citep{2011arXiv1106.4487W} related to evolution strategies) is a general
optimization technique that can be used to form a differentiable bound
on the optima of a non-differentiable function. Given a function $f$ to minimize, VO
is based on the simple fact that
\begin{equation}
    \min_{\mathbf{c} \in {\cal C}} f(\mathbf{c}) \leq \mathbb{E}_{\mathbf{c} \sim q_\psi(\mathbf{c})} [f(\mathbf{c})] = U(\psi),
\end{equation}
where $q_\psi$ is a proposal distribution with parameters $\psi$ over input values $\mathbf{c}$.
That is, the minimum of a set of function values is always less than or equal
to any of their average. Provided that the proposal is flexible enough, the parameters $\psi$
can be updated to place its mass arbitrarily tight around the optimum $\mathbf{c}^* = \min_{\mathbf{c} \in {\cal C}} f(\mathbf{c})$.

Under mild restrictions outlined in  \citep{2012arXiv1212.4507S}, the bound $U(\psi)$ is differentiable, and using the log-likelihood trick it comes:
\begin{align}\label{eqn:vo-grad}
    \nabla_\psi U(\psi) &= \nabla_\psi \mathbb{E}_{\mathbf{c} \sim q_\psi(\mathbf{c})} [f(\mathbf{c})] \nonumber \\
    &= \nabla_\psi \int f(\mathbf{c})  q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \int f(\mathbf{c}) \nabla_\psi q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \int \left[ f(\mathbf{c}) \nabla_\psi \log q_\psi(\mathbf{c}) \right]  q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \mathbb{E}_{\mathbf{c} \sim q_\psi(\mathbf{c})} [f(\mathbf{c}) \nabla_\psi \log q_\psi(\mathbf{c})]
\end{align}
Effectively, this means that provided that the score function $\nabla_\psi \log q_\psi(\mathbf{c})$ of the proposal
is known and that one can evaluate $f(\mathbf{\mathbf{c}})$ for any $\mathbf{c}$, then
one can construct empirical estimates of Eqn.~\ref{eqn:vo-grad}, which can
in turn be used to perform stochastic gradient descent (or a variant thereof)
in order to minimize $U(\psi)$.

>>>>

Following \citep{2012arXiv1212.4507S}, variational optimization (VO) is a general
optimization technique that can be used to form a differentiable bound
on the optima of a non-differentiable function. Given a function $f$ to minimize, VO
is based on the simple fact that
\begin{equation}
    \min_{\mathbf{c} \in {\cal C}} f(\mathbf{c}) \leq \mathbb{E}_{\mathbf{c} \sim q_\psi(\mathbf{c})} [f(\mathbf{c})] = U(\psi),
\end{equation}
where $q_\psi$ is a proposal distribution with parameters $\psi$ over input values $\mathbf{c}$.
That is, the minimum of a set of function values is always less than or equal
to any of their average. Provided that the proposal is flexible enough, the parameters $\psi$
can be updated to place its mass arbitrarily tight around the optimum $\mathbf{c}^* = \min_{\mathbf{c} \in {\cal C}} f(\mathbf{c})$.

Under mild restrictions outlined in  \citep{2012arXiv1212.4507S}, the bound $U(\psi)$ is differentiable, and using the log-likelihood trick it comes:
\begin{align}\label{eqn:vo-grad}
    \nabla_\psi U(\psi) &= \nabla_\psi \mathbb{E}_{\mathbf{c} \sim q_\psi(\mathbf{c})} [f(\mathbf{c})] \nonumber \\
    &= \nabla_\psi \int f(\mathbf{c})  q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \int f(\mathbf{c}) \nabla_\psi q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \int \left[ f(\mathbf{c}) \nabla_\psi \log q_\psi(\mathbf{c}) \right]  q_\psi(\mathbf{c})  d\mathbf{c} \nonumber \\
    &= \mathbb{E}_{\mathbf{c} \sim q_\psi(\mathbf{c})} [f(\mathbf{c}) \nabla_\psi \log q_\psi(\mathbf{c})]
\end{align}
Effectively, this means that provided that the score function $\nabla_\psi \log q_\psi(\mathbf{c})$ of the proposal
is known and that one can evaluate $f(\mathbf{\mathbf{c}})$ for any $\mathbf{c}$, then
one can construct empirical estimates of Eqn.~\ref{eqn:vo-grad}, which can
in turn be used to perform stochastic gradient descent (or a variant thereof)
in order to minimize $U(\psi)$.
\glnote{Consider instead natural evolution strategies, which should be more stable?}
\glnote{Rewrite in terms of NES!!}
